---
title: "Task 4 Analysis"
author: "Nick Cunnington"
date: "23/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(readxl)
library(here)
```

```{r, include=FALSE}
here::here()
candy <- read_csv(here("clean_data/candy_clean.csv"), col_types = 
                    cols(gender = col_character(), country = col_character()))
```

<br>

## Project Overview ##

### Introduction ###

This dataset is a combination of 3 different datasets for the years of 2015, 2016 and 2017 that have been joined together.  It details a survey taken of people's favourite candy types, along with some age, gender and location information.

### Assumputions ###

*During the cleaning of the data set some assumptions had to be made:*

* There was a wide variety of ages so I removed any entries for ages older that 99 
* There was no country or gender information recorded in the 2015 dataset, so any analysis requiring that information should be considered carefully as any query from 2015 will be returned as NA and may skew any result when consider all years combined



<br>

# Insights #

### What is the total number of candy ratings given across the three years? ###

```{r}
candy %>%
  filter(!is.na(rating)) %>%
  nrow()
```

<br>

### What was the average age of people who are going out trick or treating and the average age of people who are not going trick or treating? ###


```{r}
candy %>%
  group_by(going_out) %>%
  summarise(avg_age = mean(age, na.rm = TRUE))
```


<br>

### For each of the joy, despair and meh, which candy bar received the most of these ratings? ###

```{r}
candy %>%
  filter(!is.na(rating)) %>%
  group_by(rating, candy) %>%
  summarise(count = n()) %>%
  filter(count == max(count))
```

<br>

### 4.How many people rated Starburst as despair? ###

```{r}
candy %>%
  filter(candy == "starburst") %>%
  filter(rating == "despair") %>%
  summarise(id_count = n_distinct(person_id))
```

<br>

### What was the most popular candy bar for each gender? ###

*Recoding ratings to be a number for easier caculation*
```{r}
candy <- candy %>%
  mutate(rating_numeric = recode(rating, joy = 1, despair = -1, meh = 0))
```


```{r}
candy %>%
  group_by(gender, candy) %>%
  summarise(avg_rating = mean(rating_numeric, na.rm = TRUE)) %>%
  filter(avg_rating == max(avg_rating))
```

<br>

### What was the most popular candy each the US, UK, Canada and other countries? ###

```{r}
candy %>%
  mutate(country = if_else(country %in% c("united states", "canada", "united kingdom"), 
                           country, "other")) %>%
  group_by(country, candy) %>%
  summarise(avg_rating = mean(rating_numeric, na.rm = TRUE)) %>%
  filter(avg_rating == max(avg_rating))
```

<br>

### What was the msot popular candys in each year? ###

```{r}
candy %>%
  group_by(year, candy) %>%
  summarise(avg_rating = mean(rating_numeric, na.rm = TRUE)) %>%
  filter(avg_rating == max(avg_rating))
```

