---
title: "task1_analysis"
author: "Nick Cunnington"
date: "20/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(here)
here::here()
clean_data <- read_rds(here("clean_data/clean_data.rds"))
```
<br>

# Tasks #

### 1. Who had the longest long jump seen in the data? ###

```{r}
clean_data %>%
  select(names, long_jump, competition) %>%
  arrange(desc(long_jump)) %>%
  head(1)
```

<br>

### 2. What was the average 100m time in each competition? ###

```{r}
clean_data %>%
  group_by(competition) %>%
  summarise(average_100m = mean(x100m))
```

<br>

### 3. Who had the highest total points across both competitions? ###

```{r}
clean_data %>%
  select(names, points) %>%
  group_by(names) %>%
  summarise(total_points = sum(points)) %>%
  arrange(desc(total_points)) %>%
  head(1)
  
```
<br>

### 4. What was the shot-put scores for the top three competitors in each competition? ###

**Decastar Competition**
```{r}
shotput_decastar <- clean_data %>%
  select(names, shot_put, competition) %>%
  filter(competition == "Decastar") %>%
  arrange(desc(shot_put)) %>%
  head(3)
shotput_decastar
```


**OlympicG Competition**
```{r}
shotput_olympicg <- clean_data %>%
  select(names, shot_put, competition) %>%
  filter(competition == "OlympicG") %>%
  arrange(desc(shot_put)) %>%
  head(3)
shotput_olympicg
```

<br>

### 5. What was the average points for competitors who ran the 400m in less than 50 seconds vs those that ran the 400m in more than 50 seconds? ###

**Under 50 seconds**
```{r}
x400m_under50 <- clean_data %>%
  select(names, x400m, points) %>%
  filter(x400m < 50) %>%
  summarise(average_poins = mean(points))
round(x400m_under50)

```

**Over 50 seconds**
```{r}
x400m_over50 <- clean_data %>%
  select(names, x400m, points) %>%
  filter(x400m > 50) %>%
  summarise(averge_points = mean(points))
round(x400m_over50)
  
```

